---
title: "Plotting Enzymatic Data"
author: "Brady Johnston"
date: "27/08/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require('ggplot2', character.only=T, quietly=T)) {
  install.packages('ggplot2')
  library('ggplot2', character.only=T)
}
if (!require('reshape2', character.only=T, quietly=T)) {
  install.packages('reshape2')
  library('reshape2', character.only=T)
}
if (!require('ggthemes', character.only=T, quietly=T)) {
  install.packages('ggthemes')
  library('ggthemes', character.only=T)
}
if (!require('ggpubr', character.only=T, quietly=T)) {
  install.packages('ggpubr')
  library('ggpubr', character.only=T)
}
if (!require('egg', character.only=T, quietly=T)) {
  install.packages('egg')
  library('egg', character.only=T)
}
```

## Plotting enzyme data.

In this example we're going to be plotting data from multiple enzyme experiments and using some `facet_wrap()` functions to arrange it all nicely. 

```{r read data}
enzdata <- read.csv("./enzyme.csv") 
enzdata
```

## Melt Data

To make it easier to work with, we need to melt the data into a more workable format. 

```{r melt}
melt.enz <- melt(enzdata, id.vars = c("Sub"))
head(melt.enz)
```

## Start plotting
Here is the fun part where we can start plotting. 

```{r}
plt <- ggplot(melt.enz, aes(Sub, value)) + geom_point()

plt
```

We can colour-code by adding in `colour=variable` into our `aes()` 

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) + geom_point()

plt
```

Time to add some `x` and `y` labels, a title, and make the plot square.

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) +
  geom_point() +
  labs(title = "Enzyme Plots") + 
  theme_base() +
  theme(aspect.ratio = 1/1) + labs(tag="A") +
  xlab("[Substrate] nM") + ylab("Velocity µM/min")

plt
```

A good way to represent multiple experiments is by having side-by side data. We can do that with the `facet_wrap()` and telling it to separate based on variable. 

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) +
  geom_point() + facet_wrap(~variable) + 
  labs(title = "Enzyme Plots") + 
  theme_base() +
  theme(aspect.ratio = 1/1) + 
  xlab("[Substrate] nM") + ylab("Velocity µM/min")

plt
```

We don't need the legend anymore, so we can give `legend.position` a `null` variabel that will tell it to not render. 

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) +
  geom_point() + facet_wrap(~variable) +
  labs(title = "Enzyme Plots") + 
  theme_base() +
  theme(aspect.ratio = 1/1, legend.position = "") + 
  xlab("[Substrate] nM") + ylab("Velocity µM/min")

plt
```

If your figure is destined for a journal or something similar where you will letter the individual plots, you can re-label with with `tag_facet(~variable)`

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) +
  geom_point() + facet_wrap(~variable) +
  labs(title = "Enzyme Plots") + 
  theme_base() +
  theme(aspect.ratio = 1/1, legend.position = "") + 
  xlab("[Substrate] nM") + ylab("Velocity µM/min")

tag_facet(plt)
```

I like to have the data in the background so that each plot can be compared to it. To do so, we'll create a `geom_point()` but with `null` information for what it is being faceted on.

To do that we'll use `geom_point(data = transform(melt.enz, variable = NULL), colour = "grey95")` before the `geom_point()` so that the coloured dots are plotted after, and thus _on top of_ the grey dots. 

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) + 
  geom_point(data = transform(melt.enz, variable = NULL), colour = "grey90") +
  geom_point() + facet_wrap(~variable) +
  labs(title = "Enzyme Plots") + 
  theme_base() +
  theme(aspect.ratio = 1/1, legend.position = "") + 
  xlab("[Substrate] nM") + ylab("Velocity µM/min")

tag_facet(plt)
```


Finally, to add a line to these graphs we can use a function using the Michaelis-Menton equation for plotting the velocity. 

### Michaelis-Menton Function
```
enzfit <- function(num) {
  geom_smooth(method = "nls", 
              method.args = list(formula = y ~ Vmax * x / (Km + x), 
                                 start = list(Vmax = 50, Km = 0.2)),
              se = F, size = 0.5, 
              data = subset(melt.enz, variable==paste("Exp.", num, sep = "")), colour = "black")
}
```

### Adding the line to each of the 6 plots.

```
plt + enzfit(1)
head(melt.enz)
for(i in 1:6) {
  assign("plt", plt + enzfit(i), envir = .GlobalEnv)
}
```
### Final code to generate plots from the data.

```{r}
plt <- ggplot(melt.enz, aes(Sub, value, colour=variable)) + 
  geom_point(data = transform(melt.enz, variable = NULL), colour = "grey90") +
  geom_point() + facet_wrap(~variable) +
  labs(title = "Enzyme Plots") + 
  theme_base() +
  theme(aspect.ratio = 1/1, legend.position = "") + 
  xlab("[Substrate] nM") + ylab("Velocity µM/min")

enzfit <- function(num) {
  geom_smooth(method = "nls", 
              method.args = list(formula = y ~ Vmax * x / (Km + x), 
                                 start = list(Vmax = 50, Km = 0.2)),
              se = F, size = 0.5, 
              data = subset(melt.enz, variable==paste("Exp.", num, sep = "")))
}

for(i in 1:6) {
  assign("plt", plt + enzfit(i), envir = .GlobalEnv)
}

tag_facet(plt)
```